SPANG Template Library: Making SPARQL Queries FAIR

### 2020-05-23 paper.tex
v1.4

Contributions of this work:
we provide a SPARQL meta data required for constucting SPARQL queries.
we constructed template engines for SPARQL queries.
we constructed controlled SPARQL queries for available SPARQL endpoints.

@@

\begin{figure}
\center
\includegraphics[width=1.0\textwidth]{spang_top.png}
\caption{SPANG web site}
\label{fig:spang_top}
\end{figure}

@@

\subsection{Findability of SPARQL queries}
The first step is using data to find SPARQL queries. Metadata should be easy to find for both humans and computers. Machine-readable metadata are essential for automatic discovery of resources, so this is essential component.
% \begin{itemize}
%     \item Assign global identifier
%     \item Rich metadata
%     \item Include the identifier of the data they describe
%     \item Registered and indexed
% \end{itemize}

@@

\subsection{Accessibility to SPARQL queries}
Once the user finds the required data, the user needs to know how can they be accessed, possibly including authentication and authorisation.
% \begin{itemize}
%     \item Retrievable by identifier
%     \item Metadata are accessible
% \end{itemize}

@@

\subsection{Interoperability of SPARQL queries}
The ultimate goal of FAIR is to optimise the reuse of data. To achieve this, metadata should be well-described so that they can be replicated and/or combined in different settings.
% \begin{itemize}
%     \item Richly described with a plurality of accurate and relevant attributes
%     \begin{itemize}
%         \item released with a clear and accessible data usage license
%         \item associated with detailed provenance
%         \item meet domain-relevant community standards
%     \end{itemize}
% \end{itemize}

@@

\subsection{Reusability of SPARQL queries}
Reusability of SPARQL is not a new topic. Various activity is conducted to construct libraries of SPARQL. It is considered as acitivities to increase the reusability of SPARQL queries. However, no standard to increase the reusability is proposed. In the specifiation of SPARQL, reusability is not a centroal topic. However, it is an central resource in a Semantic Web. If it is a component of Semantic Web comprised of RDF/OWL/SPARQL, it is reasonable to represent it in RDF framework. We propose a way to represent SPARQL in the RDF framework.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%
\section{SPANG Library}
%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Specifications and Implementation Details}
%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsection{Using in JavaScript applications}
SPANG is now implemented in JavaScript, so the SPARQL queries can be executed in JavaScript code across the internet. To avid CORS retriction, it can access CDN services such as jsDelivr~\cite{jsdelivr} as shown in Figure~\ref{fig:jsdelivr}.

@@

\subsection{SPANG command line client}
As an interface the the SPANG library, we constructed a command-line interface.

@@

\subsection{SPANG command: calling libraries}

@@

\subsection{SPANG command: combinatorial execution}

>>>>

Contributions of this work is 
1) we provide a SPARQL meta data required for constucting SPARQL queries;
2) we constructed template engines for SPARQL queries; and
3) we constructed controlled SPARQL queries for available SPARQL endpoints.

@@

\begin{figure}
\center
\includegraphics[width=1.0\textwidth]{spang_top.png}
\caption{SPANG server top page}
\label{fig:spang_top}
\end{figure}

@@

\subsection{Findability}
The first step is using data to find SPARQL queries. Metadata should be easy to find for both humans and computers. Machine-readable metadata are essential for automatic discovery of resources, so this is essential component.
% \begin{itemize}
%     \item Assign global identifier
%     \item Rich metadata
%     \item Include the identifier of the data they describe
%     \item Registered and indexed
% \end{itemize}

@@

\subsection{Accessibility}
Once the user finds the required data, the user needs to know how can they be accessed, possibly including authentication and authorisation.
% \begin{itemize}
%     \item Retrievable by identifier
%     \item Metadata are accessible
% \end{itemize}

@@

\subsection{Interoperability}
The ultimate goal of FAIR is to optimise the reuse of data. To achieve this, metadata should be well-described so that they can be replicated and/or combined in different settings.
% \begin{itemize}
%     \item Richly described with a plurality of accurate and relevant attributes
%     \begin{itemize}
%         \item released with a clear and accessible data usage license
%         \item associated with detailed provenance
%         \item meet domain-relevant community standards
%     \end{itemize}
% \end{itemize}

@@

\subsection{Reusability}
Reusability of SPARQL is not a new topic. Various activity is conducted to construct libraries of SPARQL. It is considered as acitivities to increase the reusability of SPARQL queries. However, no standard to increase the reusability is proposed. In the specifiation of SPARQL, reusability is not a centroal topic. However, it is an central resource in a Semantic Web. If it is a component of Semantic Web comprised of RDF/OWL/SPARQL, it is reasonable to represent it in RDF framework. We propose a way to represent SPARQL in the RDF framework.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%
\section{SPANG Library and Use Cases}
%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Specifications}
%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsection{JavaScript}
SPANG is now implemented in JavaScript, so the SPARQL queries can be executed in JavaScript code across the internet. To avid CORS retriction, it can access CDN services such as jsDelivr~\cite{jsdelivr} as shown in Figure~\ref{fig:jsdelivr}.

@@

\subsection{Command-line client}
As an interface the the SPANG library, we constructed a command-line interface.

@@

\subsubsection{Calling libraries}

@@

\subsubsection{Combinatorial execution}
